### 2023-05-08 chapter2/chapter2.tex
Made some minor wording changes

The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise (Fig.~\ref{fig:hierarchy}). However, we identify at least 12 clusters with a strong and consistent patterns of constraint, with many having significant associations with certain processes, functions, or compartments. For example, cluster B is strongly enriched in proteins associated with the Wnt signaling pathway, and cluster F contains many proteins involved in nuclear transport. (Selected annotations corresponding to the summaries in Fig.~\ref{fig:hierarchy} are given in Table~\ref{stable:annotations}.)  Though these preliminary results suggest the existence of high-level relationships between certain patterns of conserved features and function, further insight into the biophysical and mechanistic details that underpin these associations will rely on more in-depth analyses of specific clusters and the properties of their constituent IDRs.

@@

\subsection{Disorder is correlated with many molecular features}
Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations of a common ancestor which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Thus, the analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. We found that disorder scores have the strongest correlations with features that measure a region's overall polarity and hydrophobicity, \textit{e.g.} fraction\_disorder and hydropathy. As the formation of a hydrophobic core is a major driving force in protein folding, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of intrinsic disorder.

@@

The GO annotation enrichment analysis indicates that proteins involved in DNA repair and extracellular structures contain a disproportionate number of regions whose disorder scores are rapidly evolving. Because the significance tests were not corrected for multiple testing or controlled for their false discovery rate, we caution against over-interpreting this result and instead consider it as a hypothesis for further investigation. In general, however, these regions with rapidly evolving disorder scores may correspond to molecular recognition features (MoRFs), which are modules in IDRs that undergo a disorder-to-order transition on binding their targets~\cite{Mohan2006, vanderLee2014}. Because MoRFs already exist on the boundary between disorder and structure, small changes in the biophysical properties of these regions may have disproportionate effects on their structural ensembles. Thus, the most variable disorder scores may reflect instances where a mutation triggered a ``phase transition'' between largely structured or disordered native states.

>>>>

The resulting heatmap reveals patterns of similar signatures interspersed among a high background of noise (Fig.~\ref{fig:hierarchy}). However, we identify at least 12 clusters with a strong and consistent patterns of constraint, with many having significant associations with certain processes, functions, or compartments. For example, cluster B is strongly enriched in proteins associated with the Wnt signaling pathway, and cluster F contains many proteins involved in nuclear transport. (Selected annotations corresponding to the summaries in Fig.~\ref{fig:hierarchy} are given in Table~\ref{stable:annotations}.) Though these preliminary results suggest the existence of a high-level relationship between certain patterns of conserved features and function, further insight into the mechanisms that underpin these associations will rely on more in-depth analyses of specific clusters and the properties of their constituent IDRs.

@@

\subsection{Disorder is correlated with many molecular features}
Though the mutations generated by evolution are not a random or exhaustive sample of sequence space, they are perturbations of a common ancestor which can reveal the relationship between a region's biophysical properties and its propensity for disorder. Thus, the analysis of feature and score contrasts is effectively a natural ``mutational scanning'' experiment. We found that disorder scores have the strongest correlations with features that measure a region's overall polarity and hydrophobicity, \textit{e.g.} fraction\_disorder and hydropathy. As the formation of a hydrophobic core is a major driving force in protein folding, this relationship is expected. However, the strength of these correlations indicate that a region's relative proportion of hydrophilic and hydrophobic residues is, to a first approximation, the largest determinant of predicted intrinsic disorder.

@@

The GO annotation enrichment analysis indicates that proteins involved in DNA repair and extracellular structures contain a disproportionate number of regions whose disorder scores are rapidly evolving. Because the significance tests were not corrected for multiple testing or controlled for their false discovery rate, we caution against over-interpreting this result and instead consider it as a hypothesis for further investigation. In general, however, these regions with rapidly evolving disorder scores may correspond to molecular recognition features (MoRFs), which are modules in IDRs that undergo a disorder-to-order transition on binding their targets~\cite{Mohan2006, vanderLee2014}. Because MoRFs already exist on the boundary between disorder and structure, small changes in the biophysical properties of these regions may have large effects on their structural ensembles. Thus, the most variable disorder scores may reflect instances where a mutation triggered a ``phase transition'' between largely structured or disordered native states.
